---
title: "TwD Project Group 2"
subtitle: "Group number: G02D01"
author: Clea Yenifer, Sutolimin Widjaja, Vincent Lavoie, Yuchi Yu
format: html
editor: visual
---

## Initialization of Libraries

```{r, message=FALSE, warning=FALSE, include=TRUE}
library(tidyverse)
library(DT)
library(lmSubsets)
library(leaps)
library(car)
library(caret)

bodyfat <- read.table("bodyfat.txt", header = TRUE, sep = "\t", check.names = TRUE)
names(bodyfat) <- tolower(names(bodyfat))
#dim(bodyfat)
#datatable(bodyfat)
```

# Data Cleaning and Analysis

## Brief EDA

-to add-

```{r}
#summary(bodyfat)
```

```{r fig.width=6, fig.height=3}
par(mfrow = c(1,2))
hist(bodyfat$pct.bf)
qqnorm(bodyfat$pct.bf)
qqline(bodyfat$pct.bf, col = "red")
par(mfrow = c(1,1))
```

```{r, message = FALSE, warning = FALSE, fig.width=6, fig.height=4}
corrplot::corrplot.mixed(cor(bodyfat),upper="ellipse", lower="number", tl.pos="lt", tl.cex = 0.6, number.cex = 0.6)
```

## Data Transformation

Height: Convert inches to cm Weight: Convert lbs to kg

```{r}
bodyfat$weight_kg <- bodyfat$weight * 0.453592
bodyfat$height_cm <- bodyfat$height * 2.54

bodyfat = bodyfat[,-c(1,4,5)]

bodyfat = bodyfat %>% filter(pct.bf > 0)
```

## Initial Model Experimentation

```{r}
ols1 = lm(pct.bf ~ .-waist, data = bodyfat)
summary(ols1)

plot(fitted(ols1), resid(ols1),
     main = "Residuals vs Fitted",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 19, col = "#4472C4",
     cex = 0.8,
     frame.plot = FALSE)
abline(h = 0, col = "#FF0000", lwd = 2, lty = 2)
lines(lowess(fitted(ols1), resid(ols1)), col = "darkgreen", lwd = 2.5)
```

```{r}
vif(ols1)
```

```{r}
ols2 = lm(pct.bf ~ age + neck + chest + abdomen + hip + thigh + knee + ankle + bicep + forearm + wrist + height_cm, data = bodyfat)
summary(ols2)
```

```{r}
vif(ols2)
```

```{r}
ols3 = lm(pct.bf ~ age + neck + abdomen + hip + knee + ankle + bicep + forearm + wrist + height_cm, data = bodyfat)
summary(ols3)
```

```{r}
vif(ols3)
```

```{r}
ols4 = lm(pct.bf ~ age + neck + abdomen + bicep + forearm + wrist + height_cm, data = bodyfat)
summary(ols4)
```

```{r}
ols5 = lm(pct.bf ~ age + abdomen + wrist + height_cm, data = bodyfat)
summary(ols5)
```

```{r}
anova(ols2, ols5)
```

# Model Comparisons

## Linear Model (from above)

```{r}
model1 = lm(pct.bf ~ age + abdomen + wrist + height_cm, data = bodyfat)
```

## Exhaustive Search

```{r}
exh_leaps = regsubsets(pct.bf ~ . -waist, data = bodyfat, nvmax = 13)
summary(exh_leaps)$outmat
```

```{r}
exh = lmSubsets(pct.bf ~ . -waist, data = bodyfat, nvmax = 13)
plot(exh)
```

```{r}
model2 = lm(pct.bf ~ age + abdomen + wrist + height_cm, data = bodyfat)
summary(model2)
```

## Backward variable selection

```{r}
n <- nrow(bodyfat)
M1 = lm(pct.bf ~ . -waist, data = bodyfat)

step.back.bic = step(M1, 
                     direction = "backward", 
                     k = log(n),
                     trace = FALSE)

#summary(step.back.bic)
round(summary(step.back.bic)$coefficients, 3)
```

## Forward variable selection

```{r}
M0 = lm(pct.bf ~ 1, data = bodyfat)

step.fwd.bic = step(M0, 
                    scope = list(lower = M0, upper = M1),
                    direction = "forward",
                    k = log(n),
                    trace = FALSE)

#summary(step.fwd.bic)
round(summary(step.fwd.bic)$coefficients, 3)
```

# Cross Validation

```{r}
set.seed(12345)
folds = 10

cv_model1 = train(pct.bf ~ age + abdomen + wrist + height_cm, data = bodyfat, method = "lm", trControl = trainControl(method="cv", number = folds, verboseIter = FALSE))
cv_model2 = train(pct.bf ~ abdomen + weight_kg + wrist, data = bodyfat, method = "lm", trControl = trainControl(method="cv", number = folds, verboseIter = FALSE))

results = resamples(list(model1 = cv_model1, model2 = cv_model2))

ggplot(results, metric = "MAE") + labs(y = "MAE")
ggplot(results, metric = "RMSE") + labs(y = "RMSE")
ggplot(results, metric = "Rsquared") + labs(y = "Rsquared")


data.frame(
  Model = c("Model 1", "Model 2"),
  RMSE = sapply(list(cv_model1, cv_model2), function(m) mean(m$resample$RMSE)),
  MAE = sapply(list(cv_model1, cv_model2), function(m) mean(m$resample$MAE)),
  Rsquared = sapply(list(cv_model1, cv_model2), function(m) mean(m$resample$Rsquared))
)
```

```{r}
final_model = lm(pct.bf ~ age + abdomen + wrist + height_cm, data = bodyfat)
```

# Tests

## Collinearity (Variance Inflation Factor)

```{r}
vif(final_model, type = "predictor")
```

## Non-Linearity of Response-Predictor Relationship

```{r}
par(mfrow = c(1, 2), bg = "white", family = "sans", mar = c(5, 5, 4, 2))

# Residuals vs Fitted
plot(fitted(final_model), resid(final_model),
     main = "Residuals vs Fitted",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 19, col = "#4472C4",
     cex = 0.8,
     frame.plot = FALSE)
abline(h = 0, col = "#FF0000", lwd = 2, lty = 2)
lines(lowess(fitted(final_model), resid(final_model)), col = "darkgreen", lwd = 2.5)

# Q-Q Plot
qqnorm(resid(final_model),
       main = "Normal Q-Q Plot",
       xlab = "Theoretical Quantiles",
       ylab = "Sample Quantiles",
       pch = 19, col = "#4472C4",
       cex = 0.8,
       frame.plot = FALSE)
qqline(resid(final_model), col = "#FF0000", lwd = 2, lty = 2)

par(mfrow = c(1, 1))
```

## Heteroskedasticity Test

```{r}
std_resid = rstandard(final_model)
sqrt_std_resid = sqrt(abs(std_resid))
plot(fitted(final_model), sqrt_std_resid,
     main = "Scale-Location Plot",
     xlab = "Fitted Values",
     ylab = "Sqrt(|Standardized Residuals|)",
     pch = 19, col = "purple")
lines(lowess(fitted(final_model), sqrt_std_resid), col = "darkgreen", lwd = 3)
rm(std_resid, sqrt_std_resid)
```

## Outliers

```{r}
plot(fitted(final_model), studres(final_model),
     main = "Studentised Residual vs Fitted Plot",
     xlab = "Fitted Values",
     ylab = "Studentised Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "darkgrey")
outliers = which(abs(studres(final_model)) > 3)
outliers
```

## Prediction Interval

```{r warning=FALSE}
predictions = predict(final_model, newdata = bodyfat, interval = "prediction", level = 0.95)

comparison = cbind(bodyfat, predictions)

ggplot(comparison, aes(x = predictions, y = pct.bf)) +
    geom_point() +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, fill = "red") +
    geom_abline(intercept = 0, slope = 1, color = "blue") +
    labs(title = "Predicted vs Actual",
         x = "Predicted", y = "Actual")
```
