---
title: "TwD Project 2"
subtitle: "Group number: G02D01"
author: Clea Yenifer, Sutolimin Widjaja, Vincent Lavoie, Yuchi Yu
format: html
editor: visual
---

## Intialization of Libraries
```{r, message=FALSE, warning=FALSE, include=TRUE}
library(tidyverse)
library(DT)
library(lmSubsets)
library(leaps)
library(car)
library(caret)

bodyfat <- read.table("bodyfat.txt", header = TRUE, sep = "\t", check.names = TRUE)
#dim(bodyfat)
#datatable(bodyfat)
```

# Data Cleaning and Analysis

## Brief EDA
-to add-

```{r}
summary(bodyfat)
```

```{r fig.width=6, fig.height=3}
par(mfrow = c(1,2))
hist(bodyfat$pct.bf)
qqnorm(bodyfat$pct.bf)
qqline(bodyfat$pct.bf, col = "red")
par(mfrow = c(1,1))
```
right skewed

## Data Transformation

Height: Convert inches to cm 
Weight: Convert lbs to kg

```{r}
names(bodyfat) <- tolower(names(bodyfat))

bodyfat$weight_kg <- bodyfat$weight * 0.453592
bodyfat$height_cm <- bodyfat$height * 2.54

bodyfat = bodyfat[,-c(1,4,5)]

bodyfat = bodyfat %>% filter(pct.bf > 0)
```

```{r, message = FALSE, warning = FALSE, fig.width=6, fig.height=4}
corrplot::corrplot.mixed(cor(bodyfat),upper="ellipse", lower="number", tl.pos="lt", tl.cex = 0.6, number.cex = 0.6)
```

## Initial Model Experimentation

```{r}
ols1 = lm(pct.bf ~ .-waist, data = bodyfat)
summary(ols1)
plot(fitted(ols1), resid(ols1),
     main = "Residual vs Fitted Plot",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "red", lwd = 2)
lines(lowess(fitted(ols1), resid(ols1)), col = "darkgreen", lwd = 2)
```
log Y isnt the right move bcz Log assumes Y can grow without limit. body fat pct is bounded 0-100 theoretically. so we 

```{r}
vif(ols1)
```
based on the vif and when we test, it actually increases adj r2, we drop weight


```{r}
ols2 = lm(pct.bf ~ age + neck + chest + log(abdomen) + hip + thigh + knee + ankle + bicep + forearm + wrist + height_cm, data = bodyfat)
summary(ols2)
plot(fitted(ols2), resid(ols2),
     main = "Residual vs Fitted Plot",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "red", lwd = 2)
lines(lowess(fitted(ols2), resid(ols2)), col = "darkgreen", lwd = 2)
```

```{r}
vif(ols2)
```

based on vif, we try removing chest, hip, thigh: they are insignificant statistically, vif high (>5), they are all strongly correl with each other and with abdomen, adj R2 increases after removing them

```{r}
ols3 = lm(pct.bf ~ age + neck + log(abdomen) + knee + ankle + bicep + forearm + wrist + height_cm, data = bodyfat)
summary(ols3)
plot(fitted(ols3), resid(ols3),
     main = "Residual vs Fitted Plot",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "red", lwd = 2)
lines(lowess(fitted(ols3), resid(ols3)), col = "darkgreen", lwd = 2)
vif(ols3)
```


# Competing Models

## Linear Model (from above)
```{r}
model1 = lm(pct.bf ~ age + neck + log(abdomen) + knee + ankle + bicep + forearm + wrist + height_cm, data = bodyfat)
```

## Exhaustive Search

```{r}
exh_leaps = regsubsets(pct.bf ~ . -waist, data = bodyfat, nvmax = 13)
summary(exh_leaps)$outmat
```

```{r}
exh = lmSubsets(pct.bf ~ . -waist, data = bodyfat, nvmax = 13)
plot(exh)
```


```{r}
model2 = lm(pct.bf ~ age + abdomen + wrist + height_cm, data = bodyfat)
summary(model2)
```

## Forward/Backward variable selection



# Cross Validation
```{r}
folds = 10
#cv_model1 = train(pct.bf ~ , method = "lm", trControl = trainControl(method="cv", number = folds, verboseIter = FALSE))
#cv_model2 = train(pct.bf ~ , method = "lm", trControl = trainControl(method="cv", number = folds, verboseIter = FALSE))
#cv_model3 = train(pct.bf ~ , method = "lm", trControl = trainControl(method="cv", number = folds, verboseIter = FALSE))
results = resamples(list(model1 = cv_model1, model2 = cv_model2, model3 = cv_model3))
ggplot(results, metric = "MAE") + labs(y = "MAE")
ggplot(results, metric = "RMSE") + labs(y = "RMSE")
ggplot(results, metric = "Rsquared") + labs(y = "Rsquared")
```


# Heteroskedasticity Test

```{r}
std_resid = rstandard(ols3)
sqrt_std_resid = sqrt(abs(std_resid))
plot(fitted(ols3), sqrt_std_resid,
     main = "Scale-Location Plot",
     xlab = "Fitted Values",
     ylab = "Sqrt(|Standardized Residuals|)",
     pch = 19, col = "purple")
lines(lowess(fitted(ols3), sqrt_std_resid), col = "darkgreen", lwd = 2)
rm(std_resid, sqrt_std_resid)
```

# Correlation of Error Terms

