---
title: "TwD Project 2"
subtitle: "Group number: G02D01"
author: Clea Yenifer, Sutolimin Widjaja, Vincent Lavoie, Yuchi Yu
format: html
editor: visual
---

## Intialization of Libraries
```{r, message=FALSE, warning=FALSE, include=TRUE}
library(tidyverse)
library(DT)
library(lmSubsets)
library(leaps)

bodyfat <- read.table("bodyfat.txt", header = TRUE, sep = "\t", check.names = TRUE)
#dim(bodyfat)
#datatable(bodyfat)
```

# Data Cleaning and Analysis

## Brief EDA
-to add-
##### why drop waist, not abdomen?
##### should we drop other variables besides waist for multicoll

## Data Transformation

Height: Convert inches to cm 
Weight: Convert lbs to kg

```{r}
names(bodyfat) <- tolower(names(bodyfat))

bodyfat$weight_kg <- bodyfat$weight * 0.453592
bodyfat$height_cm <- bodyfat$height * 2.54

bodyfat = bodyfat[,-c(1,4,5)]
```

```{r, message = FALSE, warning = FALSE, fig.width=6, fig.height=4}
corrplot::corrplot.mixed(cor(bodyfat),upper="ellipse", lower="number", tl.pos="lt", tl.cex = 0.6, number.cex = 0.6)
```

## Initial Model Experimentation

```{r}
ols1 = lm(pct.bf ~ .-waist, data = bodyfat)
#summary(ols1)
plot(fitted(ols1), resid(ols1),
     main = "Residual vs Fitted Plot",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "red", lwd = 2)
lines(lowess(fitted(ols1), resid(ols1)), col = "darkgreen", lwd = 2)
```

```{r}
bodyfat2 = bodyfat %>% filter(pct.bf > 0)
ols2 = lm(log(pct.bf) ~ .-waist, data = bodyfat2)
plot(fitted(ols2), resid(ols2),
     main = "Residual vs Fitted Plot (Linear)",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "red", lwd = 2)
lines(lowess(fitted(ols2), resid(ols2)), col = "darkgreen", lwd = 2)
```
log Y isnt the right move bcz Log assumes Y can grow without limit. body fat pct is bounded 0-100 (theoretically)

```{r}
ols3 = lm(pct.bf ~ age + neck + chest + log(abdomen) + hip + thigh + knee + ankle + bicep + forearm + wrist + height_cm + log(weight_kg), data = bodyfat)
summary(ols3)
plot(fitted(ols3), resid(ols3),
     main = "Residual vs Fitted Plot",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 19, col = "blue")
abline(h = 0, col = "red", lwd = 2)
lines(lowess(fitted(ols3), resid(ols3)), col = "darkgreen", lwd = 2)
```

# Heteroskedasticity Test

```{r}
std_resid = rstandard(ols3)
sqrt_std_resid = sqrt(abs(std_resid))
plot(fitted(ols3), sqrt_std_resid,
     main = "Scale-Location Plot (Linear)",
     xlab = "Fitted Values",
     ylab = "Sqrt(|Standardized Residuals|)",
     pch = 19, col = "purple")
lines(lowess(fitted(ols3), sqrt_std_resid), col = "darkgreen", lwd = 2)
rm(std_resid, sqrt_std_resid)
```


```{r}
rp_fas1 = lm(pct.bf ~ thigh, data = bodyfat)
rp_fas2 = lm(pct.bf ~ I(thigh^2), data = bodyfat)
rp_fas3 = lm(pct.bf ~ I(thigh^3), data = bodyfat)
rp_lgfas = lm(pct.bf ~ log(thigh), data = bodyfat)

(fas_df = data.frame(
  Metric = c("Adj. R²", "AIC", "BIC"),
  Original = c(summary(rp_fas1)$adj.r.squared, AIC(rp_fas1), BIC(rp_fas1)),
  Squared = c(summary(rp_fas2)$adj.r.squared, AIC(rp_fas2), BIC(rp_fas2)),
  Cubed = c(summary(rp_fas3)$adj.r.squared, AIC(rp_fas3), BIC(rp_fas3)),
  Logged = c(summary(rp_lgfas)$adj.r.squared, AIC(rp_lgfas), BIC(rp_lgfas))
))
```

```{r}
b1 = lm(pct.bf ~ abdomen*chest, data = bodyfat)
b2 = lm(pct.bf ~ abdomen + chest, data = bodyfat)
b3 = lm(pct.bf ~ abdomen*weight_kg, data = bodyfat)
b4 = lm(pct.bf ~ abdomen + weight_kg, data = bodyfat)
b5 = lm(pct.bf ~ hip*thigh, data = bodyfat)
b6 = lm(pct.bf ~ hip + thigh, data = bodyfat)
b7 = lm(pct.bf ~ chest*hip, data = bodyfat)
b8 = lm(pct.bf ~ chest + hip, data = bodyfat)

(b_df = data.frame(
  Metric = c("Adj. R²", "AIC", "BIC"),
  Model1 = c(summary(b1)$adj.r.squared, AIC(b1), BIC(b1)),
  Model2 = c(summary(b2)$adj.r.squared, AIC(b2), BIC(b2)),
  Model3 = c(summary(b3)$adj.r.squared, AIC(b3), BIC(b3)),
  Model4 = c(summary(b4)$adj.r.squared, AIC(b4), BIC(b4)),
  Model5 = c(summary(b5)$adj.r.squared, AIC(b5), BIC(b5)),
  Model6 = c(summary(b6)$adj.r.squared, AIC(b6), BIC(b6)),
  Model7 = c(summary(b7)$adj.r.squared, AIC(b7), BIC(b7)),
  Model8 = c(summary(b8)$adj.r.squared, AIC(b8), BIC(b8))
))
```


# Competing Models

## 1. Exhaustive Searches

```{r, echo=FALSE}

exh_leaps = regsubsets(Pct.BF ~ . - Density - Waist, data = bodyfat, nvmax = 13)
summary(exh_leaps)$outmat

```

### 1.1 ImSubsets package

```{r, echo=FALSE}
exh = lmSubsets(Pct.BF ~ . - Density - Waist, data = bodyfat, nvmax = 13)
plot(exh)

```

## 2. Linear regression of all the perimeters except density and waist

```{r, message=FALSE, warning=FALSE, echo=FALSE}
model_full <- lm(Pct.BF ~ Age + Height + Neck + Chest + Abdomen + Forearm + Wrist, data = bodyfat)

summary(model_full)
```

### 2.1 VIF

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(car)
vif(model_full)
```
