---
title: "TwD Project 2"
subtitle: "Group number: G02D00"
author: 
format: html
editor: visual
---

Insert data:

```{r, message=FALSE, include=TRUE}
library(tidyverse) 
library(DT)      

bodyfat <- read.table("bodyfat.txt", header = TRUE, sep = "\t", check.names = TRUE)
dim(bodyfat)
datatable(bodyfat)
```

# Data Cleaning and analysis

## 1. Data Conversion

Weight: Convert bls to cm Height: Convert inches to cm

```{r, message = FALSE, warning = FALSE, fig.width=12, fig.height=10}
# 1. Convert Weight (lbs to kg)
bodyfat$Weight_kg <- bodyfat$Weight * 0.453592

# 2. Convert Height (inches to cm)
bodyfat$Height_cm <- bodyfat$Height * 2.54

# 3. Bonus: Calculate BMI (Weight in kg / Height in meters squared)
bodyfat$BMI <- bodyfat$Weight_kg / (bodyfat$Height_cm / 100)^2

# View the first few rows
head(bodyfat)
```

## 2. Multiple regression

According to the multiple regression below, Body fat and Density have a very high correlation at -0.987. Additionally, Waist and Abdomen are perfectly correlated (1.00). We need to remove Density and Waist from the module to prevent multicollinearity.

```{r, message = FALSE, warning = FALSE, fig.width=12, fig.height=10}
library(GGally)
ggpairs(bodyfat)
```

# Linear Regression

## 1. Exhaustive Searches

```{r, echo=FALSE}
library(leaps)
exh_leaps = regsubsets(Pct.BF ~ . - Density - Waist, data = bodyfat, nvmax = 13)
summary(exh_leaps)$outmat

```

### 1.1 ImSubsets package

```{r, echo=FALSE}
library(lmSubsets)
exh = lmSubsets(Pct.BF ~ . - Density - Waist, data = bodyfat, nvmax = 13)
plot(exh)

```

## 2. Linear regression of all the perimeters except density and waist

```{r, message=FALSE, warning=FALSE, echo=FALSE}
model_full <- lm(Pct.BF ~ Age + Height + Neck + Chest + Abdomen + Forearm + Wrist, data = bodyfat)

summary(model_full)
```

### 2.1 VIF

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(car)
vif(model_full)
```
